Evaluating infectious disease forecasts with allocation scoring rules

### 2023-09-14 alloscore_manuscript/alloscore-application.Rnw
Update alloscore_manuscript/alloscore-application.Rnw
Co-authored-by: Evan Ray <elray1@users.noreply.github.com>

In the supplement we derive a general form of this solution by beginning with the fact that a condition for minimizing
expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.  As a consequence, at $x^{F,K}$, all the rates of change of expected loss with respect to the amount of resources allocated to a particular location $i$ must be equal:
\begin{align}
 \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = \lambda
\end{align}
for some constant $\lambda$ independent of $i$. It turns out that these derivatives can also be expressed in terms of 
the marginal forecasts $F_i$ as
\begin{align}
 \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] = L \cdot (F_i(x_i) - 1)
\end{align}
so that 
\begin{align}
F(x_i^{F,K}) = 1-\lambda/L.
\end{align}
This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need. 
It also allows us to write the components $x_i^{F,K}$ of the Bayes act as quantiles
\begin{align}
x_i^{F,K} = F_i^{-1}(1 - \lambda/L)
\end{align}
which can be substituted into the constraint equation of the allocation problem to get an equation for $\lambda$,
\begin{align}
\sum_{i=1}^N F_i^{-1}(1 - \lambda/L) = K.
\end{align}
A solution to this equation---which almost always must be approximated with a numerical method---can be substituted back into the quantile formulas
to arrive at an explicit expression for the Bayes act.
This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
We note that $\lambda = \lambda(K,F)$ and the probability level $1-\lambda/L$ depend on both the constraint level $K$ defining the decision problem and on the forecast $F$. This is quite different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.

>>>>

In the supplement we derive a general form of this solution by beginning with the fact that a condition for minimizing
expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.  As a consequence, at $x^{F,K}$, all the rates of change of expected loss with respect to the amount of resources allocated to a particular location $i$ must be equal:
\begin{align}
 \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = \lambda
\end{align}
for some constant $\lambda$ that is the same for all locations $i$. It turns out that these derivatives can also be expressed in terms of 
the marginal forecasts $F_i$ as
\begin{align}
 \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] = L \cdot (F_i(x_i) - 1)
\end{align}
so that 
\begin{align}
F(x_i^{F,K}) = 1-\lambda/L.
\end{align}
This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need. 
It also allows us to write the components $x_i^{F,K}$ of the Bayes act as quantiles
\begin{align}
x_i^{F,K} = F_i^{-1}(1 - \lambda/L)
\end{align}
which can be substituted into the constraint equation of the allocation problem to get an equation for $\lambda$,
\begin{align}
\sum_{i=1}^N F_i^{-1}(1 - \lambda/L) = K.
\end{align}
The value of $\lambda$ that solves this equation---which almost always must be approximated with a numerical method---can be substituted back into the quantile formulas
to arrive at an explicit expression for the Bayes act.
This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
We note that $\lambda = \lambda(K,F)$ and the probability level $1-\lambda/L$ depend on both the constraint level $K$ defining the decision problem and on the forecast $F$. This is quite different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.
